---
  title: "Brief Fitbit Report"
  pagetitle: "Fitbit Report"
  author: ""
  date: ""
  output: html_document
---

<hr class="heavy">
<h2> Participant ID: `r paste0("#", user)` </h2>
<h4> This report was created for the `r study_name` </h4>
<h4> by `r report_author` on `r format(Sys.Date(), "%B %d, %Y")` at `r format(Sys.time(), "%H:%M:%S")` </h4>
<center><img src=`r system.file("images/fitbit.logo.png", package = "iFitbit")` alt="Fitbit logo"/></center>

```{r set up, echo = FALSE, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{css}

h1.title {
  padding-top: 50px;
  font-size: 64px;
  color: #000000;
  font-weight: bold;
  text-align: center;
  vertical-align: middle;
}

h1 {
  padding-top: 50px;
  font-size: 64px;
  color: #000000;
  font-weight: bold;
  text-align: center;
  vertical-align: middle;
}

h2 {
  background: #B2B9CF;
  padding-top: 10px;
  padding-bottom: 10px;
  font-size: 36px;
  color: #000000;
  font-weight: bold;
  text-align: center;
  vertical-align: middle;
}

h3 {
  font-size: 24px;
  padding-top: 10px;
  padding-bottom: 10px;
  color: #536878;
  text-align: center;
  font-weight: bold;
}

h4 {
  font-size: 18px;
  color: #536878;
  text-align: center;
}

p {
  font-size: 18px;
  text-align: center;
}

hr.heavy {
  width: 100%;
  height: 10px;
  background: #536878;
  margin-top: 20px;
  margin-bottom: 20px;
}

hr.light {
  width: 100%;
  height: 5px;
  background: #536878;
  margin-top: 10px;
  margin-bottom: 10px;
}

```

```{r load libraries, include=FALSE}

library(ggplot2)
lastSynced <- as.POSIXct(device$lastSyncTime, format = "%Y-%m-%dT%H:%M:%OS", tz = "America/Chicago")

```

<p> This report was generated using R and the public <a href= "https://dev.fitbit.com/build/reference/web-api/"> Fitbit Web API </a> that can be used to retrieve Fitbit user data collected by Fitbit devices. Information that can be accessed via the <a href= "https://dev.fitbit.com/build/reference/web-api/"> Fitbit Web API </a> includes data related to <a href = "https://dev.fitbit.com/build/reference/web-api/activity/">activities</a>, <a href = "https://dev.fitbit.com/build/reference/web-api/body/">body</a>, <a href = "https://dev.fitbit.com/build/reference/web-api/devices/">devices</a>, <a href = "https://dev.fitbit.com/build/reference/web-api/friends/">friends</a>, <a href = "https://dev.fitbit.com/build/reference/web-api/heartrate-timeseries/">heart rate</a>, <a href = "https://dev.fitbit.com/build/reference/web-api/nutrition/">nutrition</a>, <a href = "https://dev.fitbit.com/build/reference/web-api/sleep/">sleep</a>, and <a href = "https://dev.fitbit.com/build/reference/web-api/user/">user</a> information. </p>

<p style="font-size:24px;color:#8B0000;font-weight:bold;"> Fitbit synced on `r format(lastSynced, "%B %d, %Y")` at  `r format(lastSynced, "%H:%M:%S")` </p>

<hr class="light">
<h3> Activity </h3>

<p> Here are two figures that report daily minutes of fairly and very active physical activity (left) and daily steps (right) from the past 7 days. The exact values of fairly and very active physical activity minutes and steps are reported as text at the top of each figure. The horizontal dotted line is the mean of activity or steps for the past 7 days. Activity data for the entirety of the intervention is reported under the plots. This table is interactive and can be filtered by date. </p><br>

```{r steps and mvpa seven and twenty-eight day average, fig.width=18, fig.height=4}

footprint.img <-  magick::image_read(system.file("images/footprint.jpeg", package = "iFitbit"))
activities$date <- lubridate::as_date(activities$date)
activities <- dplyr::mutate_if(activities, is.character, as.numeric)

findActivityAverages <- function(days){
  
  findMean <- function(variable, days) {
    activities %>% 
      dplyr::arrange(dplyr::desc(date)) %>%
      dplyr::top_n(days, date) %>%
      dplyr::mutate(mvpa = minutesFairlyActive + minutesVeryActive,
             pa = minutesLightlyActive + minutesFairlyActive + minutesVeryActive) %>%
      dplyr::summarise_at(variable, .funs = function(x) round(mean(x, na.rm = TRUE))) %>%
      purrr::as_vector() %>%
      unname()
  }
  
  ggplot() +
    geom_rect(aes(xmin = 0, xmax = 500, ymin = 0, ymax = 500), fill = "white", color = "black", size = 2) +
    annotation_raster(footprint.img, xmin = 340, xmax = 490, ymin = 240, ymax = 490) +
    annotate("text", x = 475, y = 60, label = paste(days, "Day"), hjust = 1, size = 14, fontface = "bold", color = "#002A3A") +
    annotate("text", x = 10, y = 450, label = "AVERAGES", hjust = 0, size = 14, fontface = "bold", color = "#002A3A") +
    annotate("text", x = 10, y = 350, label = paste(findMean("steps", days), "steps"), hjust = 0, size = 12) +
    annotate("text", x = 10, y = 250, label = paste("MVPA:", findMean("mvpa", days), "min./day"), hjust = 0, size = 12) +
    annotate("text", x = 10, y = 150, label = paste("Total PA:", findMean("pa", days), "min./day"), hjust = 0, size = 12) +
    theme(panel.background = element_blank(), axis.text = element_blank(),
          axis.ticks = element_blank(), axis.title = element_blank())
}
  
gridExtra::grid.arrange(
  findActivityAverages(7),
  findActivityAverages(14),
  findActivityAverages(28),
  nrow = 1
)


```

```{r fitbit mvpa plotly plot, fig.width=9, fig.height=5}

p.steps <- activities %>%
  dplyr::mutate(Category = ifelse(steps >= mean(steps, na.rm = TRUE), "Above Mean", "Below Mean")) %>%
  dplyr::rename("Date" = "date", "Steps" = "steps") %>%
  ggplot(aes(y = Steps)) +
  geom_hline(aes(yintercept = mean(Steps)), color = "#000000", linetype = "dashed") +
  geom_col(aes(x = Date, fill = Category), color = "#000000") +
  theme_minimal() + 
  scale_fill_manual(values = c("#002A3A", "#bf3626")) +
  scale_x_date(date_breaks = "1 day", date_labels = "%b %d") +
  theme(legend.position = "none", 
        plot.title = element_blank(), 
        axis.title = element_blank(), 
        panel.background = element_rect(color = "white"),
        panel.grid = element_blank(),
        axis.line = element_line(color = "black", 1))


p.steps.plotly <- plotly::ggplotly(p.steps, dynamicTicks = FALSE) %>%
  plotly::layout(showlegend = FALSE, xaxis = list(rangeslider = list(visible = TRUE),
                                          rangeselector = list(
                                            buttons = list(
                                              list(count=7, label="1w", step="day", stepmode="backward"),
                                              list(count=1, label="1m", step="month", stepmode="backward"),
                                              list(label="ALL", step="all")))))

p.mvpa <- activities %>%
  dplyr::mutate(MVPA = minutesFairlyActive+minutesVeryActive,
         Category = ifelse(MVPA >= mean(MVPA, na.rm = TRUE), "Above Mean", "Below Mean")) %>%
  dplyr::rename("Date" = "date") %>%
  ggplot(aes(y = MVPA)) +
  geom_hline(aes(yintercept = mean(MVPA)), color = "#000000", linetype = "dashed") +
  geom_col(aes(x = Date, fill = Category), color = "#000000") +
  theme_minimal() +
  scale_fill_manual(values = c("#002A3A", "#bf3626")) +
  scale_x_date(date_breaks = "1 day", date_labels = "%b %d") +
  theme(legend.position = "none", 
        plot.title = element_blank(),
        axis.title = element_blank(), 
        panel.background = element_rect(color = "white"),
        panel.grid = element_blank(),
        axis.line = element_line(color = "black", 1))

p.mvpa.plotly <- plotly::ggplotly(p.mvpa, dynamicTicks = FALSE) %>%
  plotly::layout(showlegend = FALSE, xaxis = list(rangeslider = list(visible = TRUE),
                                          rangeselector = list(
                                            buttons = list(
                                              list(count=7, label="1w", step="day", stepmode="backward"),
                                              list(count=1, label="1m", step="month", stepmode="backward"),
                                              list(label="ALL", step="all")))))

annotations = list(list(x = 0.2, y = 1.0,  text = "Steps", 
                        xref = "paper", yref = "paper", xanchor = "center", 
                        yanchor = "bottom", showarrow = FALSE),  
                   list(x = 0.8, y = 1, text = "MVPA", xref = "paper", 
                        yref = "paper", xanchor = "center", yanchor = "bottom", showarrow = FALSE))

plotly::subplot(p.steps.plotly, p.mvpa.plotly) %>% plotly::layout(annotations = annotations)
  

```

```{r reactable table}

activities %>%
  dplyr::arrange(dplyr::desc(date)) %>%
  dplyr::mutate(distance = round(distance, 2)) %>%
  dplyr::select(date, steps, minutesSedentary, minutesLightlyActive, minutesFairlyActive, minutesVeryActive,
         activityCalories, calories, distance, elevation, floors) %>%
  reactable::reactable(columns = list(
            date = reactable::colDef(name = "Date", filterable = TRUE, align = "center", minWidth = 125),
            steps = reactable::colDef(name = "Steps", align = "center"),
            minutesSedentary = reactable::colDef(name = "Sedentary", align = "center", minWidth = 110),
            minutesLightlyActive = reactable::colDef(name = "Lightly Active", align = "center"),
            minutesFairlyActive = reactable::colDef(name = "Fairly Active", align = "center"),
            minutesVeryActive = reactable::colDef(name = "Very Active", align = "center"),
            activityCalories = reactable::colDef(name = "Activity", align = "center"),
            calories = reactable::colDef(name = "Daily", align = "center"), 
            distance = reactable::colDef(name = "Distance", align = "center"), 
            elevation = reactable::colDef(name = "Elevation", align = "center"),
            floors = reactable::colDef(name = "Floors", align = "center")
            ),
          columnGroups = list(
            reactable::colGroup(name="Physical Activity", 
                                columns = c( "steps", "minutesSedentary", "minutesLightlyActive", 
                                             "minutesFairlyActive", "minutesVeryActive"), 
                                align = "center"),
            reactable::colGroup(name="Calories", 
                                columns = c("activityCalories", "calories"), 
                                align = "center")
            ),
          showPageSizeOptions = TRUE,
          pageSizeOptions = c(7, 14, 21, 28, 35),
          defaultPageSize = 7,
          searchable = TRUE,
          striped = TRUE,
          highlight = TRUE,
          bordered = TRUE,
          width = 920,
          defaultColDef = reactable::colDef(vAlign = "center", headerVAlign = "center"),
          theme = reactable::reactableTheme(
            borderColor = "#002A3A",
            borderWidth = 2,
            stripedColor = "#b2bfc3",
            highlightColor = "#7f949c",
            cellPadding = "8px 12px",
            style = list(fontFamily = "-apple-system, BlinkMacSystemFont, Segoe UI, Helvetica, Arial, sans-serif"),
            searchInputStyle = list(width = "100%")
            )
          )

```

<hr class="light">
<h3> Exercise Log </h3> 

```{r reactable table for recent workout sessions}

exercise_log %>%
  dplyr::arrange(dplyr::desc(date)) %>%
  dplyr::select(date, time, type, duration, steps, calories, hr) %>%
  reactable::reactable(columns = list(
            date = reactable::colDef(name = "Date", filterable = TRUE, align = "center"),
            time = reactable::colDef(name = "Time", align = "center"),
            type = reactable::colDef(name = "Type", align = "center"),
            duration = reactable::colDef(name = "Duration (minutes)", align = "center"),
            steps = reactable::colDef(name = "Steps", align = "center"),
            calories = reactable::colDef(name = "Calories", align = "center"),
            hr = reactable::colDef(name = "Heart Rate (bpm)", align = "center")
            ),
          showPageSizeOptions = TRUE,
          pageSizeOptions = c(5, 10, 15, 20, 25),
          defaultPageSize = 7,
          searchable = TRUE,
          striped = TRUE,
          highlight = TRUE,
          bordered = TRUE,
          width = 920,
          defaultColDef = reactable::colDef(vAlign = "center", headerVAlign = "center"),
          theme = reactable::reactableTheme(
            borderColor = "#002A3A",
            borderWidth = 2,
            stripedColor = "#b2bfc3",
            highlightColor = "#7f949c",
            cellPadding = "8px 12px",
            style = list(fontFamily = "-apple-system, BlinkMacSystemFont, Segoe UI, Helvetica, Arial, sans-serif"),
            searchInputStyle = list(width = "100%")
            )
          )

```



